### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
I'm almost done and happy about this Chapter 4. 38 pages, it's slightly too long.

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.65\linewidth]{system_model1.eps}
    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
    \label{fig:41:system_model1}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.45]{protocol.eps}
    \caption{The considered time-frequency slotted protocol. Each frame is composed by a fix-duration uplink slot in which the end-devices transmit their packets. If a packet is well received, the base station replies by transmitting an \emph{Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

@@

\paragraph{MAB model}
%
We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
Then, the end-users starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
Before sending another message (\emph{i.e.}, at time $t+1$), the end-user knows if it received or not this \emph{ACK} message.
%
For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
The goal of the end-user is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
as it is usually done in MAB problems.

@@

This section presents three different policies that will be used to assess the efficiency of the learning algorithms presented in the next section.
The first one is naive but can be used in practice, while the two others are very efficient but require full knowledge on the system (\ie, an oracle) and are thus unpractical.

@@

In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $i$, at that time slot.
The probability of successful transmission is computed as follows, because the $S_i$ static devices in each channel $i$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$ :
\begin{equation}
    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{K} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/K, \text{by uniform choice}}
\end{equation}

@@

We investigate in this section the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of everything, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, in order to avoid any signaling overhead.} is possible.
We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.

@@

Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem :

\begin{subequations}\label{eq:41:prob}
\begin{align}
    % \begin{cases}
    \underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
    \text{such that}\;\; & \sum_{i=1}^{K} D_i = D, \label{eq:41:eqCstr}\\
    & D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
    % \end{cases}
\end{align}
\end{subequations}

@@

Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $i$ :
%
$\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < K$, and $\widehat{D_{K}} = D - \sum_{i=1}^{K - 1} \widehat{D_i}$.

@@

In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices : they are \iid{} with mean $\mu_i = 1 - p S_i$.
Moreover, dynamic devices are assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
The smart devices try to learn the arms with highest means, while coordinating to choose different arms, \ie, avoid collisions in their choice, in a decentralized manner.
However, in this model it is assumed that the multiple agents can know that they experienced a collision with another agent, which is non-realistic for our problem at stake, as our model of smart device cannot do sensing nor differentiate collisions between smart and non-smart devices.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.65]{perf_learning.eps}
    \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices.}
    \label{fig:41:perf_learning}
\end{figure}

@@

The simulation code used for the experiments in Section~\ref{sub:41:numericalResults} is for MATLAB or GNU Octave,
and is open-sourced under the MIT License, at\\
\href{https://Bitbucket.org/scee_ietr/rl_slotted_iot_networks}{\texttt{Bitbucket.org/scee\_ietr/rl\_slotted\_iot\_networks}}.

>>>>

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.70\linewidth]{system_model1.eps}
    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
    \label{fig:41:system_model1}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.40]{protocol.eps}
    \caption{The considered time-frequency slotted protocol. Each frame is composed by a fix-duration uplink slot in which the end-devices transmit their packets. If a packet is well received, the base station replies by transmitting an \emph{Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

@@

\paragraph{MAB model}

We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
Then, the end-users starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
Before sending another message (\emph{i.e.}, at time $t+1$), the end-user knows if it received or not this \emph{ACK} message.
%
For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
The goal of the end-user is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
as it is usually done in MAB problems.

@@

We present three different policies that will be used to assess the efficiency of the learning algorithms presented in the next Subsection.
The first one is naive but can be used in practice, while the two others are very efficient but require full knowledge on the system (\ie, an oracle) and are thus unpractical.

@@

In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $i$, at that time slot.
The probability of successful transmission is computed as follows, because the $S_i$ static devices in each channel $i$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$:
\begin{equation}
    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{K} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/K, \text{by uniform choice}}
\end{equation}

@@

We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of everything, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, in order to avoid any signaling overhead.} is possible.
We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.

@@

Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem:
%
\begin{subequations}\label{eq:41:prob}
\begin{align}
    % \begin{cases}
    \underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
    \text{such that}\;\; & \sum_{i=1}^{K} D_i = D, \label{eq:41:eqCstr}\\
    & D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
    % \end{cases}
\end{align}
\end{subequations}

@@

Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $i$:
%
$\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < K$, and $\widehat{D_{K}} = D - \sum_{i=1}^{K - 1} \widehat{D_i}$.

@@

In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices: they are \iid{} with mean $\mu_i = 1 - p S_i$.
Moreover, dynamic devices are assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
The smart devices try to learn the arms with highest means, while coordinating to choose different arms, \ie, avoid collisions in their choice, in a decentralized manner.
However, in this model it is assumed that the multiple agents can know that they experienced a collision with another agent, which is non-realistic for our problem at stake, as our model of smart device cannot do sensing nor differentiate collisions between smart and non-smart devices.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.60]{perf_learning.eps}
    \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices. Both algorithms achieve close to optimal performance after a reasonable learning time.}
    \label{fig:41:perf_learning}
\end{figure}

@@

The simulation code used for the experiments in Section~\ref{sub:41:numericalResults} is for MATLAB or GNU Octave,
and is open-sourced under the MIT License, at\\
\href{https://Bitbucket.org/scee_ietr/rl_slotted_iot_networks}{\texttt{Bitbucket.org/scee\_ietr/rl\_slotted\_iot\_networks}}.
It was written in collaboration with Rémi Bonnefoi, in May $2017$.

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Trying to finish chapter 4 as quickly as possible.
I'm almost done and happy about this Chapter 4. 38 pages, it's slightly too long.

We implement an IoT network the following way: one gateway, one or several intelligent (\emph{i.e.}, learning) objects, embedding the proposed solution,
and a traffic generator that emulates radio interference from many other objects.
Intelligent objects communicate with the gateway with a wireless ALOHA-based protocol, which does not require any specific overhead for the learning.
%
Similarly to the previous section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
%
This solution could be added in a straightforward and costless manner in LoRaWAN networks, just by adding this feature in some or all the devices, without any modification on the network side.

% \TODOL{TODO:

% - Clean up notations, be sure it is OK with previous Sections

% - Be sure to not repeat any motivations / maths / explanations given before

% - Be sure to give enough details about the conclusions etc
% }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Motivation for this demonstration}
\label{sub:42:motivation}

In this Section, we describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol, such as the ones employed in LPWAN.
This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
%
We use a TestBed designed in 2017 by our team SCEE \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
and where the intelligence of each object corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.

@@

The rest of this Section is organized as follows.
% The system model is introduced in Section 2. In Section 3, we describe more formally both the \UCB{} and the TS algorithms.
Our implementation is presented in Section~\ref{sub:42:implementation}, with details about the user interface, and results are given in Section~\ref{sub:42:results}.
% and Section 6 concludes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GNU Radio implementation}
\label{sub:42:implementation}

In this section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
We first describe the simplified physical layer of this demo, then we present our GNU Radio implementation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Numerical results}
\label{sec:42:results}

We compare the two algorithms described in Section~\ref{sub:42:UCB} against a uniform access algorithm, that uniformly selects its channel at random.
Three objects are compared by their mean successful communication rates, on a horizon of $2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} shows the results averaged on $10$ repetitions using the same conditions.
%
Each experiment takes about half a day,
as we make objects generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
Each object was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $[15\%, 10\%, 2\%, 1\%]$.
% They did not overlap each other, as
When facing the same stationary background traffic, we see that the learning objects are both very quickly more efficient than the naive uniform object.
We obtain an improvement in terms of successful communication rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
%
In stationary environments, both the TS and \UCB{} algorithms are very efficient and converge quickly, resulting in a very strong decrease in collisions and failed communication slots. \UCB{} is faster to learn but eventually TS gives a (slightly) better average performance.

Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
%  \cite[Fig.3]{Bonnefoi17}.
Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.

\begin{figure}[!t]
	\centering
    \includegraphics[height=8.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) suffice for the two learning objects to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) object, which stays around $40\%$ of success. Similar gains of performance were obtained in many different scenarios.}
    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
\end{figure}

@@

\paragraph{Availability of data and materials}
%
The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at
\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/\\malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.

A $6$-minute \textbf{video} showing our demonstration is at \texttt{\url{youtu.be/HospLNQhcMk}}.
It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Fig.~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It especially shows the connections between the USRP boards, the Octoclock, and the master laptop.

\paragraph{Acknowledgment}
%
We acknowledge the work of two CentraleSup{\'e}lec students,
Cl{\'e}ment Barras and Th{\'e}o Vanneuville, for their GNU Radio project in Spring $2017$,
as we took inspiration in the GNU Radio code and the discussions we had at this time.

>>>>

We implement an IoT network the following way: one gateway, one or several intelligent (\emph{i.e.}, learning) objects, embedding the proposed solution,
and a traffic generator that emulates radio interference from many other objects.
Intelligent objects communicate with the gateway with a wireless ALOHA-based protocol, which does not require any specific overhead for the learning.
%
Similarly to the previous Section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
%
This solution could be added in a straightforward and costless manner in LoRaWAN networks, just by adding this feature in some or all the devices, without any modification on the network side.

% ----------------------------------------------------------------------
\subsection{Context of this demonstration}
\label{sub:42:motivation}
% ----------------------------------------------------------------------

We describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol, such as the ones employed in LPWAN.
This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
%
We use a TestBed designed in 2017 by our team SCEE \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
and where the intelligence of each object corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.

@@

\paragraph{Outline}

The rest of this Section is organized as follows.
Our implementation is presented in Section~\ref{sub:42:implementation}, with details about the user interface, and results are given in Section~\ref{sub:42:results}.

% ----------------------------------------------------------------------
\subsection{GNU Radio implementation}
\label{sub:42:implementation}
% ----------------------------------------------------------------------

In this Section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
We first describe the simplified physical layer of this demo, then we present our GNU Radio implementation.

@@

% ----------------------------------------------------------------------
\subsection{Numerical results}
\label{sec:42:results}
% ----------------------------------------------------------------------

We compare the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} against a uniform access algorithm, that uniformly selects its channel at random.
Three objects are compared by their mean successful communication rates, on a horizon of $2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} shows the results averaged on $10$ repetitions using the same conditions.
%
Each experiment takes about half a day,
as we make objects generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
Each object was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $[15\%, 10\%, 2\%, 1\%]$.
% They did not overlap each other, as
When facing the same stationary background traffic, we see that the learning objects are both very quickly more efficient than the naive uniform object.
We obtain an improvement in terms of successful communication rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
%
In stationary environments, both the TS and \UCB{} algorithms are very efficient and converge quickly, resulting in a very strong decrease in collisions and failed communication slots. \UCB{} is faster to learn but eventually TS gives a (slightly) better average performance.

Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.

\begin{figure}[!t]
	\centering
    \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) suffice for the two learning objects to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) object, which stays around $40\%$ of success. Similar gains of performance were obtained in many different scenarios.}
    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
\end{figure}

@@

\paragraph{Availability of data and materials}

The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at
\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/\\malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.

It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.
The demonstration only requires a laptop and free tools,
as the laptop should run GNU/Linux and a distribution like Ubuntu or Debian.

\paragraph{Video}

A $6$-minute \textbf{video} showing our demonstration is available online at \texttt{\url{youtu.be/HospLNQhcMk}}.
It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Fig.~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It especially shows the connections between the USRP boards, the Octoclock, and the master laptop.

\paragraph{Special acknowledgment}

We acknowledge the work of two CentraleSup{\'e}lec students,
Cl{\'e}ment Barras and Th{\'e}o Vanneuville, for their GNU Radio project in Spring $2017$,
as we took inspiration in the GNU Radio code and the discussions we had at this time.

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
I'm almost done and happy about this Chapter 4. 38 pages, it's slightly too long.

In this section, we extend the model presented before in order to take into account the possibility for retransmission of messages after a collision.
As before, we propose and evaluate different learning strategies based on MAB algorithms.

@@

\TODOL{This section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}}

@@

The aim of this section is to assess the performance of MAB algorithms for channel selection in LPWA networks, while taking into account the impact of retransmissions on the network performance.
For this reason, several decision making strategies are applied after a first retransmission (\ie, when a collision occurs).
Proposed approach employs contextual information provided by the number of retransmissions, and implemented at each device, so that no coordination among them is needed.
Moreover, our \UCB{}-based heuristics show low complexity making them suitable for being embedded in LPWA devices.

The contributions of this section are summarized as follows:
\begin{itemize}
	\item
	Firstly, we provide a close form approximation of the radio collision probability after a first retransmission.
	By doing this, we highlight the need to develop a learning approach for channel selection upon collision.

@@

\begin{figure}[htp!]  % [htbp]
	\centering
	% \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Markov_model.eps}
	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Markov_model-eps-converted-to.pdf}
	\caption{The considered Markov model for the behavior of all the devices in the network.}
	\label{fig:43:Markov_model}
\end{figure}
%

@@

\begin{figure}[htp!]  % [htbp]
	\centering
	% \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Approximation_m10.eps}
	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Approximation_m10-eps-converted-to.pdf}
	\caption{Our proposed approximation for the probability of collision at the second transmission. It is more precise for smaller values of $N$.}
	\label{fig:43:Approximation_m10}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/ResultsUCB.eps}
	\caption{
		Comparison among the exposed heuristics for the retransmission: \textcolor{blue}{Only \UCB}, \textcolor{cyan}{Random}, \textcolor{purple}{\UCB}, \textcolor{green}{$K$ \UCB}, and \textcolor{red}{Delayed \UCB}.
		The usage of the same learning policy for transmissions and retransmission is named \textcolor{blue}{Only \UCB{}},
		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as \textbf{no \UCB{}}.
		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.
	}
	\label{fig:43:mainExperiment1}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/ResultsUCB2.eps}
	\caption{
		Second scenario: learning helps a lot (a gain of $30\%$ in terms of collision probability), and learning to retransmit smartly is needed.
		We observe that the \textcolor{cyan}{random retransmission} achieves poor performance compared to the others.
	}
	\label{fig:43:mainExperiment2}
\end{figure}

@@

\paragraph{Note on the simulation code}
%
The source code (MATLAB or Octave) used for the simulations and the figures of this section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in July and August $2018$.

% % ----------------------------------------------------------------------
% \bibliographystyle{ieeetr}
% \bibliography{IEEE_WCNC__2019__Paper__BMBM}

>>>>

\graphicspath{{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/}}

In this Section, we extend the model presented before in order to take into account the possibility for retransmission of messages after a collision.
As before, we propose and evaluate different learning strategies based on MAB algorithms.

@@

% \TODOL{This Section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}}

@@

The aim of this Section is to assess the performance of MAB algorithms for channel selection in LPWA networks, while taking into account the impact of retransmissions on the network performance.
For this reason, several decision making strategies are applied after a first retransmission (\ie, when a collision occurs).
Proposed approach employs contextual information provided by the number of retransmissions, and implemented at each device, so that no coordination among them is needed.
Moreover, our \UCB{}-based heuristics show low complexity making them suitable for being embedded in LPWA devices.

The contributions of this Section are summarized as follows:
\begin{itemize}
	\item
	Firstly, we provide a close form approximation of the radio collision probability after a first retransmission.
	By doing this, we highlight the need to develop a learning approach for channel selection upon collision.

@@

\begin{figure}[htp!]  % [htbp]
	\centering
	% \includegraphics[width=0.65\linewidth]{Markov_model.eps}
	\includegraphics[width=0.70\linewidth]{Markov_model-eps-converted-to.pdf}
	\caption{The considered Markov model for the behavior of all the devices in the network.}
	\label{fig:43:Markov_model}
\end{figure}
%

@@

\begin{figure}[htp!]  % [htbp]
	\centering
	% \includegraphics[width=0.65\linewidth]{Approximation_m10.eps}
	\includegraphics[width=0.65\linewidth]{Approximation_m10-eps-converted-to.pdf}
	\caption{Our proposed approximation for the probability of collision at the second transmission. It is more precise for smaller values of $N$.}
	\label{fig:43:Approximation_m10}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.75\linewidth]{ResultsUCB.eps}
	\caption[Test of a shorter caption]{
		Comparison among the exposed heuristics for the retransmission: \textcolor{blue}{Only \UCB}, \textcolor{cyan}{Random}, \textcolor{purple}{\UCB}, \textcolor{green}{$K$ \UCB}, and \textcolor{red}{Delayed \UCB}.
		The usage of the same learning policy for transmissions and retransmission is named \textcolor{blue}{Only \UCB{}},
		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as \textbf{no \UCB{}}.
		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.
	}
	\label{fig:43:mainExperiment1}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.75\linewidth]{ResultsUCB2.eps}
	\caption{
		Second scenario: learning helps a lot (a gain of $30\%$ in terms of collision probability), and learning to retransmit smartly is needed.
		We observe that the \textcolor{cyan}{random retransmission} achieves poor performance compared to the others.
	}
	\label{fig:43:mainExperiment2}
\end{figure}

@@

\paragraph{Note on the simulation code}
%
The source code (MATLAB or Octave) used for the simulations and the figures of this Section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in July and August $2018$.

### 2019-02-14 2-Chapters/4-Chapter/chapter4.tex
I'm almost done and happy about this Chapter 4. 38 pages, it's slightly too long.

MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}, and the \UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
Many recent works show that MAB algorithms work well for real-world radio signal.
However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
In recent works like \cite{Bonnefoi17}, several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \iid.
As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.

We also present an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.

% \TODOL{Present outline of the chapter.}

@@

% ----------------------------------------------------------------------------
\section{Selfish learning for many dynamic devices with low activation probabilities in an IoT network}
\label{sec:4:firstModel}
% ----------------------------------------------------------------------

@@

@@

% ----------------------------------------------------------------------
\paragraph{Future works}
\label{sub:4:futureWorks}

>>>>

MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}.
Many recent works show that MAB algorithms work well for real-world radio signal.
However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.

We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \iid.
As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
We then validate the model with a hardware implementation, detailed in Section~\ref{sec:4:gnuradio}.
%
We conclude this Chapter by presenting an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.

@@

% ----------------------------------------------------------------------------
\section[Selfish learning for many dynamic devices in an IoT network]{Selfish learning for many dynamic devices with low activation probabilities in an IoT network}
\label{sec:4:firstModel}
% ----------------------------------------------------------------------

@@

\subsection*{Conclusions}

@@

% ----------------------------------------------------------------------
\subsection*{Future works}
\label{sub:4:futureWorks}
